---
title: "Linear Algebra in R Review"
author: "Nick Hass"
date: "1/8/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# 1.
```{r}
A.elements = c(8,6,6,8)
A <- matrix(A.elements, nrow = 2, ncol = 2)
B.elements = c(1, 3, 8, 5)
B <- matrix(B.elements, nrow = 2, ncol = 2)
```

### a. 
#### A + B
```{r}
A + B
```

### b. 
#### A - B
```{r}
A - B
```

### c. 
#### AB
```{r}
A %*% B
```

### d. 
#### BA
```{r}
B %*% A
```

# 2.
```{r}
A.elements = c(3,3,1,7)
A <- matrix(A.elements, nrow = 2, ncol = 2)
B.elements = c(2, 4, 2, 7, 10, 3)
B <- matrix(B.elements, nrow = 2, ncol = 3)
```

### a. 
#### A + B
```{r}
try(A + B)
```

Non-conformable

### b. 
#### A - B
```{r}
try(A - C)
```

Non-conformable

### c. 
#### AB
```{r}
A %*% B
```

### d. 
#### BA
```{r}
try(B %*% A)
```

Non-conformable

# 3.
```{r}
A = matrix(c(4, 11, 3, 2), nrow = 2, ncol = 2)
B = matrix(c(2, 8, 8, 12), nrow = 2, ncol = 2)
C = matrix(c(1,1,2, 1,3,5, 2,5,8), nrow = 3, ncol = 3)
D = matrix(c(4, 2, 4, 4), nrow = 2, ncol = 2)
```
```{r}
t(A)
t(B)
t(C)
t(D)
```
```{r}
isSymmetric(A)
isSymmetric(B)
isSymmetric(C)
isSymmetric(D)
```

A is not symmetric.

B is symmetric.

C is symmetric.

D is not symmetric.

# 4. 
### a. 
#### A inverse
```{r}
solve(A)
```

### b. 
#### B inverse
```{r}
solve(B)
```

### c. 
#### (AB) inverse
```{r}
solve(A %*% B)
```

### d. 
#### B inverse * A inverse and verify that (AB)inverse = B inverse * A inverse
```{r}
solve(A %*% B)
solve(B) %*% solve(A)
```

They are equal matrices.

# 5. 
#### Find the determinant of each of the following
```{r}
A = matrix(c(4, 11, 3, 2), nrow = 2, ncol = 2)
B = matrix(c(2,5,1,6, 8,333,1,10, 4,1,7,423, 0,0,0,0), nrow = 4, ncol = 4)
C = matrix(c(1,1,2, 1,3,5, 2,5,8), nrow = 3, ncol = 3)
# Do C transpose and verify that the det(C transpose) == det(c)
D = matrix(c(4,0,0,2), ncol = 2, nrow = 2)
# Do D inverse and verify that the det(D inverse) == 1/det(D)

det(A)
det(B)
det(C)
det(t(C))
det(D)
det(solve(D))
1/det(D)
```
Determinant of A: -25

Determinant of B: 0

Determinant of C: -1

Determinant of C': -1

$\text{|C| equals |C'|}$

Determinant of D: 0.125

Determinant of D inverse: 0.125

$\text{|D| equals |D}^{-1}|$

# 6. 
```{r}
A = matrix(c(7,-3,9,3, 2,4,3,1, 5,0,6,2, 8,2,5,1), nrow = 4, ncol = 4)
A
A11 <- A[1:2, 1:3]
A12 <- A[1:2, 4]
A21 <- A[3:4, 1:3]
A22 <- A[3:4, 4]

A11
A12
A21
A22
```

